Comparing binsegRcpp with other implementations of binary segmentation

### 2022-03-31 article.Rnw
comparison table

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

@@

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006}, 
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

@@

\begin{table}[t!]
\centering
\begin{tabular}{lllp{7.4cm}}
\hline
Type           & Distribution & Method   & Description \\ \hline
GLM            & Poisson      & ML       & Poisson regression: classical GLM,
                                           estimated by maximum likelihood (ML) \\
               &              & Quasi    & ``Quasi-Poisson regression'':
                                           same mean function, estimated by
                                           quasi-ML (QML) or equivalently
                                           generalized estimating equations (GEE),
                                           inference adjustment via estimated
                                           dispersion parameter \\
               &              & Adjusted & ``Adjusted Poisson regression'':
                                           same mean function, estimated by
                                           QML/GEE, inference adjustment via
                                           sandwich covariances\\
               & NB           & ML       & NB regression: extended GLM,
                                           estimated by ML including additional
                                           shape parameter \\ \hline
Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
                                           hurdle Poisson \\
               & NB           & ML       & Zero-inflated NB (ZINB),
                                           hurdle NB \\ \hline
\end{tabular}
\caption{\label{tab:overview} Overview of various count regression models. The
table is usually placed at the top of the page (\texttt{[t!]}), centered
(\texttt{centering}), has a caption below the table, column headers and captions
are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

>>>>

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern,amsmath,amssymb}

@@

@@

\begin{table}[t!]
\centering
\begin{tabular}{l|ccccc}
  package & binsegRcpp & changepoint & wbs & fpop & ruptures \\
  function & binseg & cpt.mean & sbs & multiBinSeg & Binseg \\
%   > ruptures$version$version
% [1] "1.1.6"
% > sapply(c("binsegRcpp", "changepoint", "wbs", "fpop"), function(p)paste(packageVersion(p)))
%  binsegRcpp changepoint         wbs        fpop 
% "2022.3.29"     "2.2.3"       "1.4" "2019.8.26" 
  version & 2022.3.29 & 2.2.3 & 1.4 &2019.8.26 & 1.1.6\\
\hline
  weights & yes & no & no & no & no \\
  max segs & yes & yes & no & yes & yes \\
  dim & one & one & one & multi & multi \\
  correct & yes & no & yes & yes & no \\
  losses & C++ & C & L2 & L2 & Python \\
  % all propose some loss computation.
  storage & STL multiset & arrays & recursion & heap & LRU cache \\
  space & $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$ & $O(S)$ \\
  cumsum & yes & yes & yes & yes & no \\
  best & $O(N\log N)$ & $O(N^3)$ & $O(N\log N)$ & $O(N\log N)$ & $>O(N\log N)$ \\
  worst & $O(N^2)$ & $O(N^3)$ & $O(N^2)$ & $O(N^2)$ & $O(N^2)$ \\
  CV & yes & no & no & no & no\\
  params & yes & largest & no & no & no \\
\end{tabular}
\caption{\label{tab:overview} TODO}
\end{table}

\paragraph{Loss computation.} We run each algorithm on \code{N.data}
points up to \code{max.segments =
  max.changes+1}. \code{binsegRcpp::binseg} result is a list which
contains a data table with \code{max.segments} rows and column
\code{loss} that is the square loss.  \code{changepoint::cpt.mean}
result is a list of class \code{cpt.range} with method \code{logLik}
which returns the square loss of one of the models. \code{wbs::sbs}
result is a list which contains a data frame with \code{N.data-1} rows
and \code{CUSUM} and \code{min.th} columns
TODO.\code{fpop::multiBinSeg} result is a list with element
\code{J.est}, which is a vector of \code{max.changes} square loss
decrease values.  \code{ruptures.Binseg.predict} result is a vector of
segment ends for one model size, which can be passed to
\code{sum_of_costs} method to compute the square loss.

Three packages have implemented the normal change in mean and variance
model. \code{binsegRcpp::binseg} loss values are the normal negative
log likelihood (NLL). The \code{changepoint::logLik} function returns two
times the NLL. The \code{ruptures.Binseg}
loss is related via this equation,

\begin{equation}
  \text{NLL} = (\text{rupturesLoss} + N[1+\log(2\pi)])/2
\end{equation}
